Subsidiary Recursion in Coq

### 2022-01-13 paper/paper.tex
minor

This paper is about a more expressive form of terminating recursion,
called \textbf{subsidiary recursion}.  While performing an outer
recursion on some input \verb|x|, one may initiate an inner recursion
on \verb|x| (or possibly some of its subdata), preserving the
possibility of further invocations of the outer recursive function.
Let us see a simple example.  The function \verb|wordsBy| (from
\verb|Data.List.Extra|) breaks a list into its maximal sublists whose
elements do not satisfy a predicate \verb|p|.  For example,
\verb|wordsBy isSpace " good day "| returns \verb|["good","day"]|; so
\verb|wordsBy isSpace| has the same behavior as \verb|words| (from
\verb|Data.List|).  Code is in Figure~\ref{fig:wordsBy}, using
also \verb|break| (\verb|Data.List|). Recall that
\verb|break p| is equivalent to \verb|span (not . p)|: it splits its
input into the maximal prefix whose elements do not satisfy \verb|p|,
and the following suffix.  The first recursive call,
\verb|wordsBy p tl|, is structural.  But in the second, we invoke
\verb|wordsBy p| on a value obtained from another recursion, namely
\verb|span|.  This is not allowed under structural termination, but
will be permitted by subsidiary recursion as derived below.

\begin{figure}
\begin{verbatim}
wordsBy :: (a -> Bool) -> [a] -> [[a]]
wordsBy p [] = []
wordsBy p (hd:tl) =
  if p hd
  then wordsBy p tl 
  else let (w,z) = break p tl in
        (hd:w) : wordsBy p z
\end{verbatim}
\caption{Haskell code for \texttt{wordsBy}, demonstrating subsidiary recursion}
\label{fig:wordsBy}
\end{figure}

@@

\subsection{Run-length encoding}

>>>>

This paper is about a more expressive form of terminating recursion,
called \textbf{subsidiary recursion}.  While performing an outer
recursion on some input \verb|x|, one may initiate an inner recursion
on \verb|x| (or possibly some of its subdata), preserving the
possibility of further invocations of the outer recursive function.
Let us see a simple example.  The function \verb|wordsBy| (from
\verb|Data.List.Extra|) breaks a list into its maximal sublists whose
elements do not satisfy a predicate \verb|p|.  For example,
\verb|wordsBy isSpace " good day "| returns \verb|["good","day"]|; so
\verb|wordsBy isSpace| has the same behavior as \verb|words| (from
\verb|Data.List|).  Code is in Figure~\ref{fig:wordsBy}.  The first
recursive call, \verb|wordsBy p tl|, is structural.  But in the
second, we invoke \verb|wordsBy p| on a value obtained from another
recursion, namely \verb|span|.  This is not allowed under structural
termination, but will be permitted by subsidiary recursion as derived
below.

\begin{figure}
\begin{verbatim}
wordsBy :: (a -> Bool) -> [a] -> [[a]]
wordsBy p [] = []
wordsBy p (hd:tl) =
  if p hd
  then wordsBy p tl 
  else let (w,z) = span (not . p) tl in
        (hd:w) : wordsBy p z
\end{verbatim}
\caption{Haskell code for \texttt{wordsBy}, demonstrating subsidiary recursion}
\label{fig:wordsBy}
\end{figure}

@@

\subsection{Run-length encoding (\texttt{Rle.v})}
